Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys

### 2023-07-12 SDSS-2023/testing-charts-accuracy.qmd
default to html

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: '1,2'               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Brett
    first-init: B.
    last: Berger
    marker: 2
    footnote-id: 2
  - first: Carl
    first-init: C.
    last: Camp
    marker: 3
  - first: David
    first-init: D.
    last: Dodge
    marker: 3
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
  - content: Another footnote.
    id: 2
affiliation:
  - institution: University of Achievement
    prefix: Department A        # optional
    country: Country A
    marker: 1
  - institution: Institution B
    country: Country B
    marker: 2
  - institution: Institution C
    country: Country C
    marker: 3
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
format:
  pdf:
    template: template-jdsart.cls.tex
    include-in-header: 
      - text: |
          \usepackage{amsfonts,amsmath,amssymb,amsthm}
          \usepackage{booktabs}
          \usepackage{graphicx}
          \usepackage{longtable}
          \usepackage{hyperref}
      - file: edits.tex
    cite-method: natbib
    keep-tex: true
---

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "50%",
                      fig.height = 5,
                      fig.width = 5,
                      dpi = 72#, # we can set the resolution up to 300 later.
                      #cache = TRUE) 
)
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

@@

```{r out.width='\\textwidth'}
#| label: fig-aligned-unaligned
#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e. A and C are of identical size, as are B and D. While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
knitr::include_graphics(c("images/round5-modified.PNG", "images/round3-modified.PNG"))
```

@@

```{r}
#| label: fig-tasks-12
#| fig-cap: "The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. In both instances, the tile on the right is (very slightly) larger."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
#| out-width: 100%
#| fig-asp: 1
knitr::include_graphics("../SDSS-2023/images/vertical-ab.PNG")
knitr::include_graphics("../SDSS-2023/images/vertical-cd.PNG")
```

>>>>

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: National Opinion Research Center (NORC)
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
format:
  html: default
  pdf:
    template: template-jdsart.cls.tex
    include-in-header: 
      - text: |
          \usepackage{amsfonts,amsmath,amssymb,amsthm}
          \usepackage{booktabs}
          \usepackage{graphicx}
          \usepackage{longtable}
          \usepackage{hyperref}
      - file: edits.tex
    cite-method: natbib
    keep-tex: true
    keep-md: true
---

\hh{hello?}

\kr{Kiegan, this is your editing color :) - you can choose a different color in edits.tex}

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "100%",
                      fig.height = 5,
                      fig.width = 5,
                      dev = "png", dev.args = list(type = "cairo-png")
#                      dpi = 72#, # we can set the resolution up to 300 later.
                      #cache = TRUE) 
)

library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

@@

```{r out.width='\\textwidth'}
#| label: fig-aligned-unaligned
#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e. A and C are of identical size, as are B and D. While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
knitr::include_graphics("images/round5-modified.PNG")
knitr::include_graphics("images/round3-modified.PNG")
```

@@

```{r}
#| label: fig-tasks-12
#| fig-cap: "The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. In both instances, the tile on the right is (very slightly) larger."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
#| out-width: 100%
#| fig-asp: 1
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
```

### 2023-07-12 analysis/analysis.qmd
default to html

---
title: "Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys'"
format:
  jasa-html: default
  jasa-pdf:
    include-in-header:
      - edits.tex
      - authors.tex
    keep-tex: true  
    journal:
      blinded: false
date: last-modified
bibliography: references.bib 
---

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "50%",
                      dpi = 72, # we can set the resolution up to 300 later.
                      cache = TRUE) 
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

library(readxl)
library(lubridate)
library(patchwork)
library(survey)
#library(ggmosaic)
library(ggpcp)
library(broom)
library(latex2exp)
library(gt) # grammar of tables

@@

```{r combine_rounds_1_and_2}
# all weights are scaled to add to 1 first, then lambda is used to 
# get the convex weight assignment, 
# then the weights are scaled to the effective sample size
round_12 <- round1_adj %>% 
            mutate(weight = lambda*weight/sum(weight)) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
                   ab_certain, cd_certain, weight) %>%
    rbind(round2_adj %>% 
            mutate(weight = (1-lambda)*weight/sum(weight)) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
                   ab_certain, cd_certain, weight))

@@

The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B. Nominally, the differences between the size of the marked elements can be expressed as multiples in $\delta = 0.8$, with $B = A + 2\delta = C + 3 \delta = D + 4 \delta$. A difference of $\delta = 0.8$ corresponds to about a $3^\degree$ difference for the angle of a wedge. We would expect that an increase in the difference in the values increases the accuracy in the responses. This is mitigated by the distance between wedges: more distance (should) lead to a decrease in accuracy. Similarly, if pieces have an axis or a line of reference in common, we would expect an increase in accuracy.

@@

```{r model-67, warning = FALSE}
# same model as for plot 
logit_67_correct <- svyglm(correct~task_labelled-1, family=quasibinomial(), design=survey_67)

>>>>

---
title: "Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys'"
format:
  jasa-html: default
  jasa-pdf:
    keep-tex: true  
    keep-md: true
    journal:
      blinded: false
date: last-modified
bibliography: references.bib 
---

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "50%",
                      dpi = 72, # we can set the resolution up to 300 later.
                      cache = FALSE) 
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

library(readxl)
library(lubridate)
library(patchwork)
library(survey)
#library(ggmosaic)
library(ggpcp)
library(broom)
library(latex2exp)
library(gt) # grammar of tables
library(multcomp)

@@

```{r combine_rounds_1_and_2}
# all weights are scaled to add to 1 first, then lambda is used to 
# get the convex weight assignment, 
# then the weights are scaled to the effective sample size
round_12 <- round1_adj %>% 
            mutate(weight = lambda*weight/sum(weight)) %>% 
            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
                   ab_certain, cd_certain, weight) %>%
    rbind(round2_adj %>% 
            mutate(weight = (1-lambda)*weight/sum(weight)) %>% 
            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
                   ab_certain, cd_certain, weight))

@@

The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B. Nominally, the differences between the size of the marked elements can be expressed as multiples in $\delta = 0.8$, with $B = A + 2\delta = C + 3 \delta = D + 4 \delta$. A difference of $\delta = 0.8$ corresponds to about a  3 degree difference for the angle of a wedge. We would expect that an increase in the difference in the values increases the accuracy in the responses. This is mitigated by the distance between wedges: more distance (should) lead to a decrease in accuracy. Similarly, if pieces have an axis or a line of reference in common, we would expect an increase in accuracy.

@@

```{r model-67, warning = FALSE, cache=TRUE}
# same model as for plot 
logit_67_correct <- svyglm(correct~task_labelled-1, family=quasibinomial(), design=survey_67)
